---
title: "tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(package1)
library(matrixStats)
library(Rcpp)
```

To use the function `k_means_cluster`:

```{r}
x = rbind(matrix(rnorm(100, sd = 0.3), ncol = 2), matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2))
k = 2
k_means_cluster(x, k)
```



Note if the data matrix is in higher than 3-dimension, it will need more implementation to visualize.
To includes plots if the input data matrix is in 2-dimensions:

```{r, echo=FALSE}
output = k_means_cluster(x, k)
plot(output$data, col = output$clusters, main = "plot of clustering on data", xlab = "output$data[,1]", 
     ylab = "output$data[,2]")
```

Same dataset if choosing a larger nstart value (will sacrifice effiency for robustness):
```{r}
output = k_means_cluster(x, k, nstart = 25)
plot(output$data, col = output$clusters, main = "plot of clustering with large nstart on data", xlab = "output$data[,1]", 
     ylab = "output$data[,2]")
```

Same dataset if choosing a smaller nstart value (will sacrifice robustness, not recommended):

```{r}
output = k_means_cluster(x, k, nstart = 1)
plot(output$data, col = output$clusters, main = "plot of clustering with small nstart on data", xlab = "output$data[,1]", 
     ylab = "output$data[,2]")
```

Another example simulated data with 3-dimensions:

```{r}
x = rbind(matrix(rnorm(150, sd = 0.3), ncol = 3), matrix(rnorm(150, mean = 1, sd = 0.3), ncol = 3))
k_means_cluster(x, k)
```

Another example simulated data with more than 2 clusters:
```{r}
x = rbind(matrix(rnorm(100, sd = 0.3), ncol = 2), matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), matrix(rnorm(100, mean = 2, sd = 0.3), ncol = 2))
k = 3
output = k_means_cluster(x, k)
plot(output$data, col = output$clusters, main = "plot of clustering on data", xlab = "output$data[,1]", 
     ylab = "output$data[,2]")
```

Same dataset if the best **k** is not chosen:
```{r}
k = 2
output = k_means_cluster(x, k)
plot(output$data, col = output$clusters, main = "plot of clustering on data", xlab = "output$data[,1]", 
     ylab = "output$data[,2]")
```

## Comparison

Note that K-Means clustering depends on initial centroids, and the cluster numbers are assigned randomly. The best way to evaluate the package is compare output centroids:

```{r}
c = kmeans(x, k)$centers
c = c[order(c[,1]),]
my_c = k_means_cluster(x, k)$centroids
all(c - my_c < 1e-5)
system.time(kmeans(x, k))
system.time(k_means_cluster(x, k)$centroids)
```
